### 2016-11-23 local-capability-machine.tex
Added text to the figures that illustrate scall

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$, $r_f$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{store $r$ $n$}] Stores $n$ to $r$ - notice the semantics do not allow an integer to be stored directly.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$. If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

@@

@@

    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, $r_f$, and $\bar{r}_{\var{args}}$.

@@

\item[\texttt{clear $r$}] Clears all cells $r$ has authority over. 

@@

$r_f$ & $c_f$ \\

@@

$r_{\var{args},n}$ & $w_{a,1}$\\

@@

\caption{Stack and register-file before the call}

@@

\begin{figure}
  \label{fig:stack-after-push}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
$c_{\var{stk}}' \rightarrow$  & $c_{\var{stk}}'$ \\
\cline{2-2}
   & $c_\pc'$ \\
\cline{2-2}
   & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
   & Restore code \\
\cline{2-2}
   & local stack\\
   & $\vdots$ \\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_{\pc}''$\\
\cline{2-2}
$r_0$  & $c_0$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
\cline{2-2}
$r_f$ & $c_f$ \\
\cline{2-2}
$r_{\var{args},1}$ & $w_{a,1}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{args},n}$ & $w_{a,1}$\\
\cline{2-2}
$r_{\var{priv},1}$ & $w_{p,1}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & $w_{p,m}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{Stack and register-file after everything has been pushed to the stack.}
\end{figure}

\begin{figure}
  \label{fig:stack-after-restrict-and-zero}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
$c_{\var{stk}}'' \rightarrow$  & $c_{\var{stk}}'$ \\
\cline{2-2}
   & $c_\pc'$ \\
\cline{2-2}
   & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
$c_0' \rightarrow$   & Restore code \\
   & $\vdots$\\
\cline{2-2}
   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_{\pc}^{(3)}$\\
\cline{2-2}
$r_0$  & $c_0'$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}''$ \\
\cline{2-2}
$r_{\var{args},1}$ & $w_{a,1}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{args},n}$ & $w_{a,1}$\\
\cline{2-2}
$r_{\var{priv},1}$ & 0\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & 0 \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file after the $c_{\var{stk}}'$ has been limited to only give authority over the empty part of the stack (the new capability is $c_{\var{stk}}''$. $c_0'$ is made from $c_{\var{stk}}'$ by setting it to point to the restore code and restricting it to a local enter-capability. }
\end{figure}

\begin{figure}
  \label{fig:stack-upon-return}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2} 
   & $?$\\
\cline{2-2} 
$c_{\var{stk}}' \rightarrow$  & $c_{\var{stk}}'$ \\
\cline{2-2}
   & $c_\pc'$ \\
\cline{2-2}
   & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
$c_0' \rightarrow$   & Restore code \\
   & $\vdots$\\
\cline{2-2}
   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_0'$\\
\cline{2-2}
$r_0$  &  ? \\
\cline{2-2}
$r_{\var{stk}}$  & ? \\
\cline{2-2}
$r_1$ & $w_1$ \\
\cline{2-2}
$r_{\var{priv},1}$ & ?\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & ? \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file upon return. }
\end{figure}

\begin{figure}
  \label{fig:stack-after-restore-code}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$c_{\var{stk}}^{(3)} \rightarrow$  & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
   & Restore code \\
   & $\vdots$\\
\cline{2-2}
   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_\pc'$\\
\cline{2-2}
$r_0$  &  ? \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}^{(3)}$ \\
\cline{2-2}
$r_1$ & $w_1$ \\
\cline{2-2}
$r_{\var{priv},1}$ & ?\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & ? \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file after clearing stack. }
\end{figure}

\begin{figure}
  \label{fig:stack-after-restore-code}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$c_{\var{stk}}^{(4)} \rightarrow$   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_\pc^{(3)}$\\
\cline{2-2}
$r_0$  &  $c_0$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}^{(4)}$ \\
\cline{2-2}
$r_1$ & $w_1$ \\
\cline{2-2}
$r_{\var{priv},1}$ & $w_{p,1}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & $w_{p,m}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file upon return. }
\end{figure}

>>>>

\begin{description}
\item[\texttt{call} $f(\bar{r}_{\var{args}},\bar{r}_{\var{priv}})$] $\bar{r}_{\var{args}}$ and $\bar{r}_{\var{priv}}$ are lists of registers. An overview of this call:
  \begin{itemize}
  \item Set up activation record
  \item Create local enter capability for activation
  \item Clear unused registers
  \item Jump
  \item Upon return: Run activation code
  \end{itemize}
A more detailed description of each of the above steps:
\begin{description}
  \item [Set up activation record]
    \begin{itemize}
    \item Run malloc to get a piece of memory with space for:
    \begin{itemize}
      \item Words in $\bar{r}_{\var{priv}}$
      \item Code return capability (opc)
      \item Activation code
    \end{itemize}
    \item Store the words in $\bar{r}_{\var{priv}}$ to the activation record. 
    \item Adjust the a copy of the current pc to point to the return address and save it to the activation record.
    \item Write the activation code to the activation record.
  \end{itemize}
  \item [Create local enter capability for activation] Make the capability for the activation record point to the beginning of the activation record and restrict it to a local-enter capability. Place this capability in $r_0$.
  \item [Clear unused registers]
    Clear all the register that are not $\pcreg$, $r_0$ or in $\bar{r}_{\var{args}}$.
  \item [Jump] Load the entry $f$ from the linking table to a free register and jump to it.
  \item [Activation code] The activation code does the following:
    \begin{itemize}
    \item Move the stored ``private'' words in to their respective $\bar{r}_{\var{priv}}$ registers.
%    \item Clear all registers but $r_1$, $\pcreg$, and $\bar{r}_{\var{priv}}$.
    \item Load the return address to $\pcreg$
    \end{itemize}
\end{description}
\item[\texttt{malloc $r$ $n$}] Calls malloc to allocates a piece of memory of size $n$. The capability will be stored in register \texttt{r}.
\item[\texttt{assert$_{\var{flag}}$ $r_1$ $r_2$}] Compares the words in register $r_1$ and $r_2$. If they are equal, then execution continues. If they are unequal, then the assertion flag named $flag$ in the flag list is set to 1 and execution halts (if no flag is specified, then the first flag in the list is set to 1).
\end{description}
Note:
\begin{itemize}
\item \texttt{call} will fail if we have local capabilities as part of the ``private'' state as it relies on a capability returned by malloc which will not be permit-write-local. Below, we introduce \texttt{scall} which can handle local capabilities in the ``private'' state.
\end{itemize}
\subsection{Stack}
Some programs will assume access to a stack which will be in part indicated by the program macros, but also in the correctness lemma. The stack is accessed through a local $\rwlx$-capability. Programs will assume that the stack resides in some register, say $r_{\var{stk}}$.

@@

Eventhough the memory is infinite, we will only use a finite part for the stack. If we have allocated too little memory for the stack, and we try to push something anyway, then the execution will fail. As we consider failing admissable, we are okay with this.

@@

    Clear all registers but $\pcreg$, $r_0$, $r_{\var{stk}}$, and $\bar{r}_{\var{args}}$.

@@

\item[\texttt{mclear $r$}] Clears all cells $r$ has authority over. 

@@

@@

$r_{\var{args},n}$ & $w_{a,n}$\\

@@

\caption{This is the first figure of 6 that illustrates how \texttt{scall} works. In this example, the call \texttt{scall $f([r_{\var{args},1},\dots,r_{\var{args},n}],[r_0,r_{\var{priv},1},\dots,r_{\var{priv},m}])$. In this example the two lists of registers are disjoined eventhough that does not have to be the case.}}

@@

\begin{figure}
  \label{fig:stack-after-push}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
$c_{\var{stk}}' \rightarrow$  & $c_{\var{stk}}'$ \\
\cline{2-2}
   & $c_\pc'$ \\
\cline{2-2}
   & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
   & Restore code \\
\cline{2-2}
   & local stack\\
   & $\vdots$ \\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_{\pc}''$\\
\cline{2-2}
$r_0$  & $c_0$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}'$ \\
\cline{2-2}
$r_{\var{args},1}$ & $w_{a,1}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{args},n}$ & $w_{a,n}$\\
\cline{2-2}
$r_{\var{priv},1}$ & $w_{p,1}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & $w_{p,m}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{Stack and register-file after the restore code, ``private'' registers (remember $r_0$ is here private.), return address ($c_\pc'$), and stack capability ($c_{\var{stk}}'$) have been pushed to the stack.}
\end{figure}

\begin{figure}
  \label{fig:stack-after-restrict-and-zero}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
$c_{\var{stk}}'' \rightarrow$  & $c_{\var{stk}}'$ \\
\cline{2-2}
   & $c_\pc'$ \\
\cline{2-2}
   & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
$c_0' \rightarrow$   & Restore code \\
   & $\vdots$\\
\cline{2-2}
   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_{\pc}^{(3)}$\\
\cline{2-2}
$r_0$  & $c_0'$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}''$ \\
\cline{2-2}
$r_{\var{args},1}$ & $w_{a,1}$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{args},n}$ & $w_{a,n}$\\
\cline{2-2}
$r_{\var{priv},1}$ & 0\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & 0 \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file after the $c_{\var{stk}}'$ has been limited to only give authority over the empty part of the stack (the new capability is $c_{\var{stk}}''$). $c_0'$ is made from $c_{\var{stk}}'$ by setting it to point to the restore code and restricting it to a local enter-capability. The ``private'' registers have been cleared.}
\end{figure}

\begin{figure}
  \label{fig:stack-upon-return}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2} 
   & $?$\\
\cline{2-2} 
$c_{\var{stk}}' \rightarrow$  & $c_{\var{stk}}'$ \\
\cline{2-2}
   & $c_\pc'$ \\
\cline{2-2}
   & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
$c_0' \rightarrow$   & Restore code \\
   & $\vdots$\\
\cline{2-2}
   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_0'$\\
\cline{2-2}
$r_0$  &  ? \\
\cline{2-2}
$r_{\var{stk}}$  & ? \\
\cline{2-2}
$r_1$ & $w_1$ \\
\cline{2-2}
$r_{\var{priv},1}$ & ?\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & ? \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file upon return from $f$. At this point we have no idea what is in the register-file apart from the $\pcreg$ which we know points to the restore code. The contents of the stack we released access to is also unknown. (Notice that we have changed the order of the registers as we are no longer interested in the argument registers. By convention we expect a return value to be in $r_1$, which is why we have named that word, but the words in the remaining non-special-purpose registers could also be considered return values.)}
\end{figure}

\begin{figure}
  \label{fig:stack-after-restore-code}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
   & ? \\
\cline{2-2}
$c_{\var{stk}}^{(3)} \rightarrow$  & $c_0$ \\
\cline{2-2}
   & $w_{p,1}$ \\
\cline{2-2}
   & $\vdots$ \\
\cline{2-2}
   & $w_{p,m}$ \\
\cline{2-2}
   & Restore code \\
   & $\vdots$\\
\cline{2-2}
   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_\pc'$\\
\cline{2-2}
$r_0$  &  ? \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}^{(3)}$ \\
\cline{2-2}
$r_1$ & $w_1$ \\
\cline{2-2}
$r_{\var{priv},1}$ & ?\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & ? \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-file after executing the restore code. The old stack capability has been restored and the  $\pcreg$-register now points to the return address in memory. }
\end{figure}

\begin{figure}
  \label{fig:stack-after-restore-code}
  \centering
  \begin{tabular}[!h]{r | >{\raggedright\arraybackslash}p{3cm} |}
\multicolumn{2}{l}{Stack} \\
\cline{2-2}
   & \\
   & $\vdots$\\
\cline{2-2}
   & 0 \\
\cline{2-2}
$c_{\var{stk}} \rightarrow$   & local stack\\
   & $\vdots$\\
\cline{2-2}
\end{tabular}
\hspace{1cm}
\begin{tabular}{r | >{\centering\arraybackslash}p{0.75cm} |}
\multicolumn{2}{r}{Register file} \\
\cline{2-2}
$\pcreg$ & $c_\pc^{(3)}$\\
\cline{2-2}
$r_0$  &  $c_0$ \\
\cline{2-2}
$r_{\var{stk}}$  & $c_{\var{stk}}$ \\
\cline{2-2}
$r_1$ & $w_1$ \\
\cline{2-2}
$r_{\var{priv},1}$ & $w_{p,1}$\\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
$r_{\var{priv},m}$ & $w_{p,m}$ \\
\cline{2-2}
& $?$ \\
\cline{2-2}
& $\vdots$ \\
\cline{2-2}
\end{tabular}
\caption{ Stack and register-register file after the clean up code has been run. The untrusted part of the stack has been cleared. The ``private'' words have been poped to their respective registers. The restore code has been poped of the stack. The stack is now in the same state as it was before the call. }
\end{figure}
